---
title: "Practical Machine Learning"
output: html_document
---
##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ??? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website [here]( http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset). 

##Data
The data for this project come from this [source](http://groupware.les.inf.puc-rio.br/har)
The training data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) and the test data [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)
 
Downloading two datasets
```{r, include= FALSE}
downloadcsv <- function(url, nastrings) {
  temp <- tempfile()
  download.file(url, temp, method = "curl")
  data <- read.csv(temp, na.strings = nastrings)
  unlink(temp)
  return(data)
}

train <- downloadcsv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", c("", "NA", "#DIV/0!"))
test <- downloadcsv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", c("", "NA", "#DIV/0!"))

```
The training data has 19622 observations and 160 features, and the distribution of the five measured stances A,B,C,D,E is:

```{r}
dim(train)
```
```{r}
table(train$classe)
```
##Preprocessing
###Partioning the training set
Separation training data into a training set and validation set.
 
```{r}
library(caret)
```
```{r}
set.seed(123456)
trainset <- createDataPartition(train$classe, p = 0.8, list = FALSE)
Training <- train[trainset, ]
Validation <- train[-trainset, ]
```
##Feature selection
Cleaning up the zero variance features, columns with missing values and descriptive fields.
```{r}
# exclude near zero variance features
nzvcol <- nearZeroVar(Training)
Training <- Training[, -nzvcol]

# exclude columns with m40% ore more missing values exclude descriptive
# columns like name etc
cntlength <- sapply(Training, function(x) {
    sum(!(is.na(x) | x == ""))
})
nullcol <- names(cntlength[cntlength < 0.6 * length(Training$classe)])
descriptcol <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
    "cvtd_timestamp", "new_window", "num_window")
excludecols <- c(descriptcol, nullcol)
Training <- Training[, !names(Training) %in% excludecols]
```
##Model Train

Random forest for classification and regression using package randomForest
```{r}
library(randomForest)

```
```{r}
rfModel <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 10)
```
##Model Validation

###Training set accurasy
```{r}
ptraining <- predict(rfModel, Training)
print(confusionMatrix(ptraining, Training$classe))
```
* Model perfome excellent against the trainig set. SO we need to do cross-validation to check if we have avaided overfittiing.

### Validation set accuracy (Out-of-Sample)
Cross validation
```{r}
pvalidation<-predict(rfModel, Validation)
print(confusionMatrix(pvalidation, Validation$classe))
```
The cross validation shows accuracy 99,5% and out-of-sample erros is 0,5%, so model performs rather good and we have avoid overfitting.

##Test set prediction

The prediction for the test set is:
```{r}
ptest<-predict(rfModel, test)
ptest
```
Submission
```{r}
answers <- as.vector(ptest)

pml_write_files = function(x) {
    n = length(x)
    for (i in 1:n) {
        filename = paste0("problem_id_", i, ".txt")
        write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
            col.names = FALSE)
    }
}

pml_write_files(answers)

```


